Toward single particle reconstruction without particle picking: Breaking the detection limit

### 2018-07-30 TEX/draft_biology.tex
Rewrote part of the appendix

\newcommand{\II}{\mathcal{I}}
\newcommand{\kk}{\textbf{k}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\mb}{\mathbf}
\newcommand{\mk}{\mathfrak}
\newcommand*\Bell{\ensuremath{\boldsymbol\ell}}

@@

\TODO{To rewrite for the discrete case}

We consider the moments of $\II$ in the limit $N\to\infty$. To take this limit, we shall assume that $N=\Omega(M^2)$, and that $\gamma = \lim_{N\to\infty}\frac{N}{M^2}\in(0,1)$ is a constant. 
Because of the separation condition Eq.~(\ref{eq:sep_cond}), if $\mb i$ is in the support of the $j$th projection, then $\mb i + \Delta\mb i$ is either in the support of the same projection or outside the support of any projection. Formally, $\mb i$ is in the support of the $j$th projection if and only if $\mb 0\leq \mb i - \mb s_j < (L,L)^T$ where the inequalities apply to each coordinate. Then because of the separation condition,

note that if $\rr\in S_i$ for some $i$, then $\rr+\Delta\rr\in S_i\cup Z$ whenever $||\Delta\rr||\leq 2L$ by the separation assumption $d> 2L$. Since $\RR^2=Z\sqcup\bigsqcup_{i=1}^NS_i$, and $\II(\rr)=0$ if $\rr\in Z$, we have for $||\Delta\rr||\leq 2L$
\[\begin{aligned} m_2[\II](\Delta \rr)  &= \frac{1}{M^2}\int_{\RR^2}\II(\rr)\II(\rr+\Delta\rr)\, d\rr\\ 
&= \frac{1}{M^2}\int_Z\II(\rr)\II(\rr+\Delta\rr)\, d\rr + \frac{1}{M^2}\sum_{i=1}^N\int_{S_i}\II(\rr)\II(\rr+\Delta\rr)\, d\rr\\ 
&= \frac{1}{M^2}\sum_{i=1}^N\int_{\RR^2}I_{\omega_i}(\rr)I_{\omega_i}(\rr+\Delta\rr)\, d\rr\\ 
&= \frac{N}{M^2}\cdot\frac{1}{N}\sum_{i=1}^NL^2m_2[I_{\omega_i}](\Delta\rr)\\
&\to \gamma\int_{\text{SO}(3)}L^2m_2[I_{\omega}](\Delta\rr)\, d\omega.\end{aligned}\]
If we set $m_2(\Delta \rr) = 0$ for $||\Delta\rr||>2L$, the above equality holds for all $\Delta\rr$, and taking its Fourier transform (with respect to $\Delta\rr$) and interchanging the Fourier integral with the integral over SO(3) (which can \emph{always} be done), we get
\[ \widehat{m_2[\II]}(\kk) \to \gamma\int_{\text{SO}(3)}L^2\widehat{m_2[I_{\omega}]}(\kk)\, d\omega = \gamma\langle|\widehat I_{\omega}(\kk)|^2\rangle_{\omega},\]
where $\langle\cdot\rangle_{\omega}$ denotes average over the orientations, so over SO(3).

Similarly, for the third moment if $\rr\in S_i$ for some $i$ then $\rr+\Delta\rr_1,\rr+\Delta\rr_2\in S_i\cup Z$, and hence we again get for $||\Delta\rr_1||\leq 2L$ and $||\Delta\rr_2||\leq 2L$ that
\[\begin{aligned} m_3[\II](\Delta \rr_1,\Delta\rr_2)  &= \frac{1}{M^2}\int_{\RR^2}\II(\rr)\II(\rr+\Delta\rr_1)\II(\rr+\Delta\rr_2)\, d\rr\\ 
&= \frac{1}{M^2}\int_Z\II(\rr)\II(\rr+\Delta\rr_1)\II(\rr+\Delta\rr_2)\, d\rr + \frac{1}{M^2}\sum_{i=1}^N\int_{S_i}\II(\rr)\II(\rr+\Delta\rr_1)\II(\rr+\Delta\rr_2)\, d\rr\\ 
&= \frac{1}{M^2}\sum_{i=1}^N\int_{\RR^2}I_{\omega_i}(\rr)I_{\omega_i}(\rr+\Delta\rr_1)I_{\omega_i}(\rr+\Delta\rr_2)\, d\rr\\ 
&= \frac{N}{M^2}\cdot\frac{1}{N}\sum_{i=1}^NL^2m_3[I_{\omega_i}](\Delta\rr_1,\Delta\rr_2)\\
&\to \gamma\int_{\text{SO}(3)}L^2m_3[I_{\omega}](\Delta\rr_1,\Delta\rr_2)\, d\omega.\end{aligned}\]
Again, setting $m_3(\Delta\rr_1,\Delta\rr_2)=0$ if either $||\Delta\rr_1||>2L$ or $||\Delta\rr_2||>2L$, we also get
\[ \widehat{m_3[\II]}(\kk_1,\kk_2) = \gamma\langle\widehat I_{\omega}(\kk_1)\widehat I_{\omega}(\kk_2)\overline{\widehat I_{\omega}(\kk_1+\kk_2)}\rangle_{\omega}.\]

\section{Steering}\label{sec:steering}
The first two moments are cheap to compute and store - $m_1$ is a scalar and $m_2$ is radially symmetric (in the limit $N\to\infty$), so it suffices to compute a single ray of it. The third moment however is infeasible to store - it has $(2L-1)^4$ entries, and in modern microscopes we can easily have $L\geq 300$. In addition, the above scheme does not average over in-plane rotations, which can improve estimation in the precense of noise (and accelerate convergence to the population moments without noise). We therefore propose the following steering procedure:

The third moment $m_3[\II](\Delta \rr_1,\Delta \rr_2)$ as defined above is compactly supported (at least numerically) with respect to each of its variables in both real and Fourier space, with support radii $2L$ and $1/2$ in real and Fourier space, respectively (see above formulas for $m_3$ in real and Fourier space). Therefore, we can expand it in a suitable product basis that is compactly supported in both real and Fourier space, e.g. Prolate Spheroidal Wave Functions (PSWFs) or Fourier-Bessel (FB):
\[ m_3[\II](\Delta \rr_1, \Delta \rr_2) = \sum_{k_1,k_2=-\infty}^{\infty}\sum_{q_1,q_2=1}^{\infty}\mathfrak{m}_3(k_1,q_1;k_2,q_2)\psi_{k_1,q_1}(\Delta\rr_1)\overline{\psi_{k_2,q_2}(\Delta\rr_2)},\]
where we conjugate the second set of basis functions for convenience, that will become apparent below (of course, the set of conjugates is also a basis, as $\overline{\psi_{k,q}}=\psi_{-k,q}$ for PSWFs and $\overline{\psi_{k,q}}=(-1)^k\psi_{-k,q}$ for FB). We shall assume that the basis $\{\psi_{k,q}(\rr)\}$ is orthonormal, so $\int_{\RR^2}\psi_{k_1,q_1}(\rr)\overline{\psi_{k_2,q_2}}(\rr) = \delta_{k_1,k_2}\delta_{q_1,q_2}$, valid for both PSWFs and FB. Then, the coefficients are given by
\[\begin{aligned} \mathfrak{m}_3(k_1,q_1;k_2,q_2) &= \int_{||\Delta\rr_1||,||\Delta\rr_2||\leq 2L}m_3[\II](\Delta\rr_1,\Delta\rr_2)\overline{\psi_{k_1,q_1}}(\Delta\rr_1)\psi_{k_2,q_2}(\Delta\rr_2)\\
&= \int_{\rr}\II(\rr)\left(\int_{||\Delta\rr_1||\leq 2L}\II(\rr+\Delta\rr_1)\overline{\psi_{k_1,q_1}}(\Delta\rr_1)\right)\left(\int_{||\Delta\rr_2||\leq 2L}\II(\rr+\Delta\rr_2)\psi_{k_2,q_2}(\Delta\rr_2)\right).\end{aligned}\]
Note that if we expand $\II(\rr+\Delta\rr)$ in $\{\psi_{k,q}(\Delta\rr)\}$ for fixed $\rr$ and $||\Delta\rr||\leq 2L$, so
\[ \II(\rr+\Delta\rr) = \sum_{k,q}a_{k,q}(\rr)\psi_{k,q}(\Delta\rr),\quad a_{k,q}(\rr)=\int_{||\Delta\rr||\leq 2L}\II(\rr+\Delta\rr)\overline{\psi_{k,q}}(\Delta\rr),\]
and use the fact that $\II$ is real, our expression becomes
\[ \mathfrak{m}_3(k_1,q_1;k_2,q_2) = \int_{\rr}\II(\rr)a_{k_1,q_1}(\rr)\overline{a_{k_2,q_2}}(\rr).\]

We further assume that all in-plane rotations of the micrograph and its reflection, or equivalently, all in-plane rotations of each such disc of radius $2L$ and its reflection, are present in our dataset. Noting that rotations and reflections commute with the Fourier transform and using the derivation of [Zhao, Landa], the expansion of the disc about $\rr$ by an angle $\alpha$ is given by
\[ \II^{\alpha,+}(\rr+\Delta\rr) = \sum_{k,q}a_{k,q}e^{-ik\alpha}\psi_{k,q}(\Delta\rr),\]
and the expansion of the reflection of that disc rotated by an angle $\alpha$ by
\[ \II^{\alpha,-}(\rr+\Delta\rr) = \sum_{k,q}\overline{a_{k,q}}e^{-ik\alpha}\psi_{k,q}(\Delta\rr),\]
where we used the fact that for real-valued images we have $a_{-k,q}=\overline{a_{k,q}}$ for both PSWFs and FB. We thus have
\[\begin{aligned} \mathfrak{m}_3(k_1,q_1;k_2,q_2) &= \int_{\rr}\II(\rr)\left(\frac{1}{4\pi}\int_0^{2\pi}\left[a_{k_1,q_1}(\rr)\overline{a_{k_2,q_2}}(\rr) + \overline{a_{k_1,q_1}}(\rr)a_{k_2,q_2}(\rr)\right]e^{-i(k_1-k_2)\alpha}d\alpha\right),\\
&= \delta_{k_1,k_2}\int_{\rr}\II(\rr)\Re\{a_{k_1,q_1}(\rr)\overline{a_{k_2,q_2}}(\rr)\}\\
&= \Re\left\{\delta_{k_1,k_2}\int_{\rr}\II(\rr)a_{k_1,q_1}(\rr)\overline{a_{k_2,q_2}}(\rr)\right\},\end{aligned}\]
Thus, the bispectrum in our steerable basis $\mathfrak{m}_3(k_1,q_1;k_2,q_2)$ is block-diagonal, and is effectively a 3-tensor.

Similarly, the power spectrum is compactly supported in space (with
support $2L-1$) and bandlimited as in Fourier space it is the average
squared magnitude of the Fourier transform of the projections, each
of which is supposedly bandlimited. Therefore, we may expand it as
\[ m_2[\II](\Delta\rr) =
\sum_{k,q}\mathfrak{m}_2(k,q)\psi_{k,q}(\Delta\rr),\]
and obtain the expansion coefficients as
\[\begin{aligned} 
\mathfrak{m}_2(k,q) &=
\int_{\Delta\rr}m_2[\II](\Delta\rr)\overline{\psi_{k,q}(\Delta\rr)}\\
&=
\int_{\rr}\II(\rr)\int_{\Delta\rr}\II(\rr+\Delta\rr)\overline{\psi_{k,q}(\Delta\rr)}\\
&= \int_{\rr}\II(\rr)a_{k,q}(\rr). \end{aligned}\]
Taking all rotations of $\II$ and its reflection, we get
\[\begin{aligned} 
\mathfrak{m}_2(k,q) &= \int_{\rr}\II(\rr)\left(\frac{1}{4\pi}\int_0^{2\pi}[a_{k,q}(\rr) +
\overline{a_{k,q}}(\rr)]e^{-ik\alpha}\, d\alpha\right)\\ 
&= \delta_{k,0} \int_{\rr}\II(\rr)a_{0,q}(\rr), \end{aligned}\]
where in the last equality we dropped the real part since both
$\II(\rr)$ and $\psi_{0,q}$ are real valued, and hence $a_{0,q}(\rr)=\int_{\Delta\rr}\II(\rr+\Delta\rr)\psi_{0,q}(\Delta\rr)$
is real as well.
Thus, the average power spectrum is effectively a vector.

\section{Connection to volume}

We derive a relation between the steered bispectrum derived in Sect.~\ref{sec:steering} to the volume, expanded in a suitable basis. Specifically, expand the Fourier-transformed volume as
\[ \widehat V(c\kk) = \sum_{\ell,m,s}a_{\ell,m,s}j_{\ell,s}(k)Y_{\ell,m}(\kk/k),\]
for $k\leq 1$ and zero otherwise, where $c$ is the assumed bandlimit, $Y_{\ell,m}$ may be taken to be either real or complex, and $j_{\ell,s}(r)$ is some radial basis function (in practice, either spherical bessel or the radial part of the 3D PSWFs as in [Lederman]). We work here with the complex spherical harmonics, given by
\[ Y_{\ell,m}(\theta,\varphi) = \sqrt{\frac{2\ell+1}{4\pi}\cdot\frac{(\ell-m)!}{(\ell+m)!}}P_{\ell}^m(\cos\theta)e^{i m\varphi},\]
where $P_{\ell}^m$ are the associated Legendre polynomials with the Condon-Shortley phase, and $j_{\ell,s}$ are normalized spherical Bessel functions given by
\[ j_{\ell, s}(k) = \frac{4}{|j_{\ell+1}(u_{\ell, s})|}j_{\ell}(2u_{\ell,s} k),\]
where $j_{\ell}$ is the spherical Bessel function of order $\ell$, $u_{\ell,s}$ is the $s$th positive zero of $j_{\ell}$, and the radius satisfies $k\in[0,1/2]$ in Fourier space since $1/2$ is the Nyquist frequency [ISBI].

Then
\[ \widehat I_{\omega}(ck,\theta) = \sum_{\ell,m,m',s}a_{\ell,m,s}Y_{\ell,m'}(\pi/2,0)D_{m',m}^{\ell}(\omega)j_{\ell,s}(k)e^{im'\theta},\]
whenever $k\leq 1$. Express this in 2D PSWFs so
\[ \widehat I_{\omega}(ck,\theta) = \sum_{N,n}b_{N,n}\psi_{N,n}(k,\theta),\]
where in papers the 2D PSWFs are defined as
\[ \psi_{N,n}(k,\theta) = \frac{1}{\sqrt{2\pi}}R_{N,n}(k)e^{ik\theta},\]
but in Boris' code the expansion is actually performed with respect to
\[ \widetilde\psi_{N,n}(k,\theta) = \frac{1}{2\sqrt{2\pi}}\alpha_{N,n}R_{N,n}(k)e^{ik\theta},\]
where $\alpha_{N,n}$ are the eigenvalues associated with $\psi_{N,n}$ - see Sect.~\ref{sec:PSWFs_props} below.

>>>>

\newcommand{\II}{\mathcal{I}}
\newcommand{\kk}{\textbf{k}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\mb}{\mathbf}
\newcommand{\mk}{\mathfrak}
\newcommand{\mc}{\mathcal}

\newcommand*\Bell{\ensuremath{\boldsymbol\ell}}

@@

We consider the moments of $\II$ in the limit $N\to\infty$. To take this limit, we shall assume that $N=\Omega(M^2)$, and that $\gamma = \lim_{N\to\infty}\frac{NL^2}{M^2}\in(0,1)$ is a constant. 
Because of the separation condition Eq.~(\ref{eq:sep_cond}), if $\mb i$ is in the support of the $j$th projection, then $\mb i + \Delta\mb i$ is either in the support of the same projection or outside the support of any projection. Formally, $\mb i$ is in the support of the $j$th projection if and only if $0\leq \mb i - \mb s_j < L$ where the inequalities apply to each coordinate. Then for $||\Delta\mb i||_{\infty}\leq L-1$ we have
\[-(L-1) \leq \mb i+\Delta\mb i-\mb s_j< 2L-1,\]
and because of the separation condition, given $\mb s_k$ we must either have $\mb s_k-\mb s_i\geq 2L-1$ in which case 
\[ \mb i + \Delta\mb i- \mb s_k \leq \mb i + \Delta\mb i- \mb s_j - (2L-1) < 0,\]
or $\mb s_k - \mb s_i \leq -(2L-1)$, in which case
\[ \mb i + \Delta\mb i - \mb s_k \geq -(L-1) + 2L-1 = L.\]
In either case, $\mb i+\Delta\mb i$ is not in the support of the $k$th particle.

Denote by $\mc S_j$ the support of the $j$th particle and by $Z=\{1,\ldots, M\}-\sqcup_j \mc S_j$ all the points outside the support of the particle. The clean micrograp $\widetilde \II = \II-\xi$ then satisfies $\widetilde \II(\mb i) = 0$ if $\mb i\in Z$, so we have
\be\begin{aligned} 
m_k[\widetilde\II](\Delta\mb i_1,\ldots, \Delta\mb i_{k-1}) &= \frac{1}{M^2}\sum_{\mb i}\widetilde\II(\mb i)\widetilde\II(\mb i + \Delta\mb i_1)\cdots \widetilde\II(\mb i + \Delta\mb i_{k-1})\\
&= \frac{1}{M^2}\sum_{j=1}^N\sum_{\mb i\in\mc S_i}\widetilde\II(\mb i)\widetilde\II(\mb i + \Delta\mb i_1)\cdots \widetilde\II(\mb i + \Delta\mb i_{k-1}) + \frac{1}{M^2}\sum_{\mb i\in Z}\widetilde\II(\mb i)\widetilde\II(\mb i + \Delta\mb i_1)\cdots \widetilde\II(\mb i + \Delta\mb i_{k-1})\\
&= \frac{NL^2}{M^2}\cdot\frac{1}{N}\sum_{j=1}^N\frac{1}{L^2}\sum_{i,j=1}^L I_{\omega_j}(\mb i)I_{\omega_j}(\mb i+\Delta\mb i_1)\cdots I_{\omega_j}(\mb i + \Delta \mb i_{k-1})\\
&= \frac{NL^2}{M^2}\frac{1}{N}\sum_{j=1}^Nm_k[I_{\omega_j}]\\
&\to \gamma\langle m_k[I_{\omega_j}]\rangle_{\omega}. \end{aligned}\ee

In the precence of noise, we get an additional bias term, which we proceed to derive. The mean is unbiased, since the noise is assumed to have zero mean. For the power spectrum, we have
\[\begin{aligned}
&m_2[\II](\Delta \mb i) =
\frac{1}{M^2}\sum_{i,j=1}^M\II(\mb i)\II(\mb
i+\Delta \mb i)\\
&= \frac{1}{M^2}\sum_{i,j=1}^M\widetilde\II(\mb i)\widetilde\II(\mb i+\Delta \mb i) + \frac{1}{M^2}\sum_{i,j=1}^M\widetilde\II(\mb i)\xi(\mb i + \Delta\mb i)\\ &+ \frac{1}{M^2}\sum_{i,j=1}^M\xi(\mb i)\widetilde\II(\mb i + \Delta\mb i) + \frac{1}{M^2}\sum_{i,j=1}^M\xi(\mb i)\xi(\mb i + \Delta\mb i). 
\end{aligned}\]
Considering the terms one-by-one, the first term is independent of the
noise, and as shown above converges to $\gamma \langle
m_2[I_{\omega}]\rangle_{\omega}$ as $N\to\infty$. The second term
satisfies
\[\begin{aligned} 
\frac{1}{M^2}\sum_{\mb i}\widetilde\II(\mb i)\xi(\mb i + \Delta\mb i) &=
\frac{NL^2}{M^2}\cdot\frac{1}{NL^2}\sum_{j, \mb i}I_{\omega_j}(\mb
i)\xi_{\omega_j}(\mb i + \Delta\mb i)\\
&\to \gamma\mathbb{E}_{\omega,\xi}[I_{\omega}\xi]\\ 
&= \gamma \mathbb{E}_{\omega}[I_{\omega}]\mathbb{E}_{\xi}[\xi] = 0, \end{aligned}\]
where $\xi_{\omega_j}(\mb i) = \xi(\mb i + \mb s_j)$ and we used the fact that the orientations $\omega$ and the noise $\xi$ are independent, and $\mathbb{E}_{\omega}[I_{\omega}] = \langle m_1[I_{\omega}]\rangle_{\omega}$. A similar argument applied to the third term shows that it
also vanishes as $N\to\infty$. 
For the fourth term,
if $\Delta\mb i \neq \mb{0}$ then since the noise is zero mean and
i.i.d. this term vanishes. If $\Delta\mb i = \mb0$ then
\[ \frac{1}{m^2}\sum_{i,j=1}^m\xi(\mb i)^2 \to \sigma^2.\]
Thus, we conclude
\[ m_2[\II](\Delta\mb i) \to \gamma\langle m_2[I_{\omega}](\Delta\mb i)\rangle_{\omega} + \sigma^2\delta(\Delta\mb i).\]
Note that the bias term $b_2 = \sigma^2\delta(\Delta\mb i)$ depends only on the variance of the noise $\sigma^2$.

For the third moments, we get 8 terms:
\[\begin{aligned} 
&m_3[\II](\Delta\mb i_1, \Delta\mb i_2) =
\underbrace{\frac{1}{M^2}\sum_{\mb i}\widetilde\II(\mb i)\widetilde\II(\mb i+\Delta\mb
	i_1)\widetilde\II(\mb i + \Delta\mb i_2)}_{(1)} +
\underbrace{\frac{1}{M^2}\sum_{\mb i}\xi(\mb i)\xi(\mb i+\Delta\mb i_1)\xi(\mb i + \Delta\mb i_2)}_{(2)}\\ 
&+ \underbrace{\frac{1}{M^2}\sum_{\mb i}\widetilde\II(\mb i)\xi(\mb i + \Delta\mb i_1)\widetilde\II(\mb i + \Delta\mb i_2)}_{(3)} +
\underbrace{\frac{1}{M^2}\sum_{\mb i}\widetilde\II(\mb i)\widetilde\II(\mb i + \Delta\mb i_1)\xi(\mb i + \Delta\mb i_2)}_{(4)}\\
&+ \underbrace{\frac{1}{M^2}\sum_{\mb i}\xi(\mb i)\widetilde\II(\mb i + \Delta\mb i_1)\widetilde\II(\mb i + \Delta\mb i_2)}_{(5)} +
\underbrace{\frac{1}{M^2}\sum_{\mb i}\widetilde\II(\mb i)\xi(\mb i + \Delta\mb i_1)\xi(\mb i + \Delta\mb i_2)}_{(6)}\\
&+ \underbrace{\frac{1}{M^2}\sum_{\mb i}\xi(\mb i)\xi(\mb i + \Delta\mb i_1)\widetilde\II(\mb i + \Delta\mb i_2)}_{(7)} +
\underbrace{\frac{1}{M^2}\sum_{\mb i}\xi(\mb i)\widetilde\II(\mb i + \Delta\mb i_1)\xi(\mb i + \Delta\mb i_2)}_{(8)}.
\end{aligned}\]
We address these terms one by one:
\begin{itemize}
	\item Term (1) is $m_3[\widetilde\II]$, shown above to converge to
	$\gamma\langle m_3[I_{\omega}]\rangle_{\omega}$.
	\item Term (2) is $m_3[\xi]$, the bispectrum of pure noise, which
	vanishes \TODO{Either cite a signal processing paper, or show ourselves}
	\item Terms (3)-(5) depend linearly on the noise and hence
	converge to zero.
	\item For term (6), if $\Delta\mb i_1\neq \Delta\mb i_2$ the term
	vanishes as then $\xi(\mb i + \Delta\mb i_1)$ and $\xi(\mb i +
	\Delta\mb i_2)$ are independent. If $\Delta\mb i_1 = \Delta\mb
	i_2$ the term becomes
	\[ \begin{aligned}
        \frac{1}{M^2}\sum_{\mb i}\widetilde\II(\mb i)\xi(\mb i + \Delta\mb i_1)^2
      &= \frac{NL^2}{M^2}\cdot\frac{1}{NL^2}\sum_{j, \mb
		i}I_{\omega_j}(\mb i)\xi_{\omega_j}(\mb i + \Delta\mb i)^2\\ 
      &\to \gamma\mathbb{E}_{\omega, \xi}[\xi^2I_{\omega}]\\
      &= \gamma\sigma^2\mathbb{E}[m_1[I_{\omega}]]\\ 
      &= \gamma\sigma^2Lm_1[\phi],
    \end{aligned}\]
	where $m_1[\phi]$ is the mean of the volume. That is because by the Fourier projection-slice theorem, the Fourier transform of the projections correspond to the restriction of the 3D Fourier transform of the volume to a slice passing through the origin, so in particular the DC components of each projection and the volume are identical, so $\sum_{i,j,k}V(i,j,k) = \sum_{i,j}I_{\omega}(i,j)$.
	\item For term (7), if $\Delta\mb i_1 \neq \mb 0$ then once again
	this term vanishes, whereas if $\Delta\mb i_2 = \mb 0$ then it becomes
	\[ \frac{1}{M^2}\sum_{\mb i}\xi(\mb i)^2\widetilde\II(\mb i + \Delta\mb
	i_2) = \frac{NL^2}{M^2}\cdot\frac{1}{NL^2}\sum_{\omega,\mb
		i}\xi_{\omega}(\mb i)^2I_{\omega}(\mb i + \Delta\mb i_2) \to
	\gamma\sigma^2Lm_1[\phi].\]
	Similarly, term (8) vanishes if $\Delta\mb i_2\neq \mb 0$ and
	converges to $\gamma\sigma^2Lm_1[\phi]$ otherwise.
\end{itemize}
Thus, we conclude that
\[ m_3[\II](\Delta\mb i_1, \Delta\mb i_2) \to \gamma\langle
m_3[I_{\omega}](\Delta\mb i_1, \Delta\mb i_2)\rangle_{\omega} +
\gamma\sigma^2Lm_1[\phi]\Big(\delta(\Delta\mb i_1 - \Delta\mb i_2) +
\delta(\Delta\mb i_1) + \delta(\Delta\mb i_2)\Big).\]
Note that in practice, we have $m_1[\II] \approx m_1[\widetilde\II] \approx \gamma
Lm_1[\phi]$ since the mean converges relatively fast to its expectation (at a rate of $1/M$), so we do not need prior knowledge of $\gamma$ to effectively
debias the bispectrum.

\section{Steering}\label{sec:steering}
The clean autocorrelations $\gamma\langle m_k[I_{\omega}]\rangle_{\omega}$ which we estimate from the micrograph are given by an average over the particule orientations. In particular, they are invariant under in-plane rotations of the projections, and can therefore be block-diagonalized in any steerable basis, a basis in which an in-plane rotation corresponds to a phase shift of the expansion coefficients - see \cite{bhamre2016denoising,landa2017steerable,zhao2016fast}. For this paper, we represent our autocorrelations using Prolate Spheroidal Wave Functions (PSWFs) $\{\psi_{k,q}\}$, which we use to include all in-plane rotations of the projections in the micrograph to compute the autocorrelations, and compute only the blocks on the diagonal - see \cite{landa2017steerable}. The PSWFs are eigenfunctions of the truncated Fourier trasnform and satisfy
\be\label{eq:PSWF_defn_eq}
\alpha_{k,q}\psi_{k,q}(\mb k) = \sum_{\mb i}\psi_{k,q}(\mb i)e^{-2\pi i (\mb i-\mb 1)\cdot(\mb k - \mb 1)}
\ee
given in polar coordinates by
\[ \psi_{k,q}(r,\varphi) = \left\{\begin{array}{ll} \frac{1}{2\sqrt{2\pi}}\alpha_{k,q}R_{k,q}(r)e^{ik\varphi}, & r\leq 1,\\ 0, & r>1,\end{array}\right. \quad k\geq 0,\ q\geq 1,\]
and are orthonormal on the unit disc.\footnote{In \cite{landa2017steerable}, PSWFs are defined as $2\psi_{k,q}/\alpha_{k,q}$, but using the above definition simplifies the formula for the expansion coefficients.}

The mean is a scalar and is already invariant under in-plane rotations of the projections in the dataset, since all projections have the same mean $Lm_1[\phi]$, as explained above. For the power spectrum, we expand it as
\[ m_2[\II](\Delta\mb i) =
\sum_{k,q}\mathfrak{m}_2(k,q)\psi_{k,q}(\Delta\mb i),\]
where the expansion is truncated according to the rule in \cite{landa2017steerable}, and obtain the expansion coefficients as
\[\begin{aligned} 
\mathfrak{m}_2(k,q) &=
\sum_{\Delta\mb i}m_2[\II](\Delta\mb i)\overline{\psi_{k,q}(\Delta\mb i)}\\
&=
\frac{1}{L^2}\sum_{\mb i}\II(\mb i)\left(\sum_{\Delta\mb i}\II(\mb i+\Delta\mb i)\overline{\psi_{k,q}(\Delta\mb i)}\right)\\
&= \frac{1}{L^2}\sum_{\mb i}\II(\mb i)a_{k,q}(\mb i), \end{aligned}\]
where we note that if we expand a patch of size $(2L-1) \times (2L-1)$ centered about $\II(\mb i)$ in PSWFs as
\[ \II(\mb i + \Delta\mb i) = \sum_{k,q}a_{k,q}(\mb i)\psi_{k,q}(\Delta\mb i),\quad ||\Delta\mb i||_{\infty}\leq L-1,\]
the expansion coefficients are given by the inner product
\[ a_{k,q}(\mb i) = \sum_{\Delta\mb i}\II(\mb i + \Delta\mb i)\overline{\psi_{k,q}(\Delta\mb i)}.\]
As shown in \cite{landa2017steerable, zhao2016fast}, the rotation of the above patch by an angle $\alpha$ is given by
\[ \II^{\alpha,+}(\mb i+\Delta\mb i) = \sum_{k,q}a_{k,q}e^{-ik\alpha}\psi_{k,q}(\Delta\mb i),\]
and the reflection of that disc rotated by an angle $\alpha$ by
\[ \II^{\alpha,-}(\mb i+\Delta\mb i) = \sum_{k,q}\overline{a_{k,q}}e^{-ik\alpha}\psi_{k,q}(\Delta\mb i),\]
where we used the fact that for real-valued images we have $a_{-k,q}=\overline{a_{k,q}}$ for both PSWFs and FB. Averaging over all rotations of the patch $\II(\mb i + \Delta\mb i)$ and its reflection we get
\[\begin{aligned} 
\mathfrak{m}_2(k,q) &= \frac{1}{L^2}\sum_{\mb i}\II(\mb i)\left(\frac{1}{4\pi}\int_0^{2\pi}[a_{k,q}(\mb i) +
\overline{a_{k,q}}(\mb i)]e^{-ik\alpha}\, d\alpha\right)\\ 
&= \delta_{k,0} \frac{1}{L^2}\sum_{\mb i}\II(\mb i)a_{0,q}(\mb i), \end{aligned}\]
where in the last equality we dropped the real part since both
$\II$ and $\psi_{0,q}$ are real valued.
Thus, the average power spectrum is effectively a vector. To estimate the clean power spectrum coefficients, we also expand the bias term in PSWFs. Writing
\[ b_2(\Delta\mb i) = \sigma^2\delta(\Delta\mb i) = \sum_{k,q}\mathfrak{b}_2(k,q)\psi_{k,q}(\Delta\mb i),\]
the coefficients are given by
\[\begin{aligned} 
\mathfrak{b}_2(k,q) &= \sigma^2\sum_{\Delta\mb i}\delta(\Delta\mb i)\overline{\psi_{k,q}(\Delta\mb i)}\\
&= \sigma^2\overline{\psi_{k,q}(\mb 0)}\\
&= \frac{\sigma^2}{\sqrt{2\pi}}R_{0,q}(0). \end{aligned}\]
We then estimate the expansion coefficients of the clean power spectrum as
\[ \widetilde{\mk m_2}(k,q) = \mk m_2(k,q) - \mk b_2(k,q).\]

We expand the bispectrum as
\[ m_3[\II](\Delta \mb i_1, \Delta \mb i_2) = \sum_{\substack{k_1,q_1\\ k_2,q_2}}\mathfrak{m}_3(k_1,q_1;k_2,q_2)\psi_{k_1,q_1}(\Delta\mb i_1)\overline{\psi_{k_2,q_2}(\Delta\mb i_2)},\]
The expansion coefficients are then given by
\[\begin{aligned} \mathfrak{m}_3(k_1,q_1;k_2,q_2) &= \sum_{\Delta\mb i_1, \Delta\mb i_2}m_3[\II](\Delta\mb i_1,\Delta\mb i_2)\overline{\psi_{k_1,q_1}(\Delta\mb i_1)}\psi_{k_2,q_2}(\Delta\mb i_2)\\
&= \frac{1}{L^2}\sum_{\mb i}\II(\mb i)\left(\sum_{\Delta\mb i_1}\II(\mb i+\Delta\mb i_1)\overline{\psi_{k_1,q_1}(\Delta\mb i_1)}\right)\left(\sum_{\Delta\mb i_2}\II(\mb i+\Delta\mb i_2)\psi_{k_2,q_2}(\Delta\mb i_2)\right)\\
&= \frac{1}{L^2}\sum_{\mb i}\II(\mb i) a_{k_1,q_1}(\mb i)\overline{a_{k_2,q_2}(\mb i)}.
\end{aligned}\]
Averaging over all rotations of $\II$ and its reflection, we obtain
\[\begin{aligned} \mathfrak{m}_3(k_1,q_1;k_2,q_2) &= \frac{1}{L^2}\sum_{\mb i}\II(\mb i)\left(\frac{1}{4\pi}\int_0^{2\pi}\left[a_{k_1,q_1}(\mb i)\overline{a_{k_2,q_2}(\mb i)} + \overline{a_{k_1,q_1}(\mb i)}a_{k_2,q_2}(\mb i)\right]e^{-i(k_1-k_2)\alpha}d\alpha\right),\\
&= \delta_{k_1,k_2}\frac{1}{L^2}\sum_{\mb i}\II(\mb i)\Re\{a_{k_1,q_1}(\mb i)\overline{a_{k_2,q_2}(\mb i)}\}.\end{aligned}\]
Thus, the bispectrum in our steerable basis $\mathfrak{m}_3(k_1,q_1;k_2,q_2)$ is block-diagonal, and is effectively a 3-tensor.
Similarly to the power spectrum, we also expand the bias term in PSWFs as
\[ b_3(\Delta\mb i_1, \Delta\mb i_2) =  \gamma\sigma^2Lm_1[\phi]\Big(\delta(\Delta\mb i_1 - \Delta\mb i_2) +
\delta(\Delta\mb i_1) + \delta(\Delta\mb i_2)\Big) = \sum_{\substack{k_1, q_1\\ k_2, q_2}} \mk 
b_3(k_1,q_1; k_2, q_2)\psi_{k_1,q_1}(\Delta\mb i_1)\overline{\psi_{k_2,q_2}(\Delta\mb i_2)},\]
where the coefficients are given by
\[ \mk b_3(k_1,q_1; k_2,q_2) = \gamma\sigma^2 m_1[\phi]\delta_{k_1,k_2}\left[\delta_{q_1,q_2} + \delta_{k_1,0}\frac{1}{2\pi}(\alpha_{0,q_1} + \alpha_{0,q_2})R_{0,q_1}(0)R_{0,q_2}(0)\right]. \]
\section{Connection to volume}

We derive a relation between the steered bispectrum derived in Sect.~\ref{sec:steering} to the volume, expanded in a suitable basis. Specifically, expand the Fourier-transformed volume as
\[ \widehat \phi(c\kk) = \sum_{\ell,m,s}a_{\ell,m,s}j_{\ell,s}(k)Y_{\ell,m}(\kk/k),\quad k\leq 1,\]
where $k=||\kk||_2$, $c$ is the assumed bandlimit, $Y_{\ell}^m$ is the complex spherical harmonics
\[ Y_{\ell,m}(\theta,\varphi) = \sqrt{\frac{2\ell+1}{4\pi}\cdot\frac{(\ell-m)!}{(\ell+m)!}}P_{\ell}^m(\cos\theta)e^{i m\varphi},\]
